Preventing a Nuclear ISIL: \\Recommendations for United States Policy

### 2015-05-11 draft.tex
Made various consistency checks.

DOS - Department of State

@@

Following the interception of SNM or weapon components, careful analysis of the composition, material structure, and relative sophistication of the physics package will allow forensic investigators to estimate the age of the device, the techniques which were utilized in order to create the physics package, and the most probable country of origin \cite{Glaser2008}. In the U.S., a number of organizations would be involved in these assessments. The National Technical Nuclear Forensics Center (NTNFC) is responsible for ensuring the nation's nuclear forensics capability.  The NTNFC is supported by the Department of Homeland Security (DHS), Department of Justice (DOJ), Federal Bureau of Investigation (FBI), Department of Defense (DoD), Department of Energy (DOE), Department of State (DOS), and the Office of the Director of National Intelligence (ODNI) \cite{DepartmentOfHomelandSecurit}.

@@

\ldots The State Department manages several programs that offer NIS scientists employment and research opportunities, so that they will be less likely to sell their knowledge to rogue nations. The International Science and Technology Center (ISTC) in Moscow and the Science and Technology Center in Ukraine (STCU), multilateral programs supported by the EU, Japan, and others, fund research in such areas as nuclear reactor safety, medicine, civil aviation, and energy production. A \enquote{Partners Program} links NIS scientists with contributing U.S. private sector companies, such as General Electric and Lockheed Martin, and academic institutes to assist their own R\&D efforts. The Civilian Research and Development Foundation (CRDF) is a private, non-profit foundation, originally funded by matching grants from DOD and George Soros. It supports U.S. collaboration and exchanges with former NIS defense scientists on civilian basic and applied research, and supports projects with commercial applications. The State Department also implements a pilot project for biological weapons scientists to \enquote{redirect} their expertise toward commercial endeavors. Among other activities, it has established a collaborative program between the Department of Agriculture's Agricultural Research Service and Russian scientists on agriculture research, and between the Department of Health and Human Services' Centers for Disease Control, National Institutes of Health, and Food and Drug Administration and Russian biomedical scientists to conduct research on infectious diseases. }

@@

Contributing Agency: DNDO, DTRA, DOS, Bechtel Nevada  \normalfont

@@

The lead U.S. agency responsible for implementing this policy will be the U.S. Department of State (DOS) supported by the U.S. Defense Threat Reduction Agency (DTRA). Preventing ISIL from acquiring a nuclear weapon, or at the very least SNM, falls under the purview of both the Bureau of Counterterrorism (CT) and the Bureau of International Security and Nonproliferation (ISN) within the DOS. Under this policy strategy, ISN is the principal actor in the DOS as they are the primary U.S. interface with the IAEA on the NPT \cite{Circular1970}. DTRA is a Combat Support Agency that supports the Combatant Commanders and Services in responding to any WMD threat. One aspect of this support is managing a research and development portfolio to develop tools and capabilities to detect and respond to WMD threats. DTRA is the agency within the U.S. government that works to verify treaties related to WMD are being carried out by all signatories, and thus will be able to support the DOS ISN in this policy \cite{DefenseThreatReductionAgency}.

@@

Lead Agency: DOT, DoD

@@

Lead Agency: DOS

@@

Lead Agency: DOS, DNDO/NTNFC

@@

The proposed policy  would strengthen the commitment to nuclear forensics by the NFAA through increased funding for the program from current \$20M 2015 budget to \$30M for 2016 \cite{Year2015}.  This money would be used for nuclear forensics-related research, to \enquote{provide steady funding streams} for nuclear engineers and physicists, to develop forensics capabilities in the area of post-detonation debris collection, and increase the number, fidelity, and participation in nuclear forensic exercises.  In addition to the increased funding, the proposed policy  would have the President, in coordination with the DOS and DoD, issue a declaratory policy regarding the U.S. response to the transfer of nuclear technology, materials, or weapons that result in a detonation of a nuclear weapon on the U.S. or one of its allies.  The proposed declaratory policy would be incorporated into the next NPR, but issued as a standalone document in the interim due to the dynamic ISIL threat.  The proposed policy wording is:

@@

\chapter{???????  Do final consistancy checks for ISIS/ISIL, US/U.S., apostrophes, SLD/SLoD, quote marks  ???????????}

@@

Intelligence agencies gather information from physical and technical observations, conversations, searches, and human operations. These sources are grouped into Signals Intelligence (SIGINT), Imagery Intelligence (IMINT), Measurement and Signature Intelligence (MASINT), Human-Source Intelligence (HUMINT), Open-Source Intelligence (OSINT) and Geospatial Intelligence (GEOINT) \cite{Intelligen}. The National Security Agency (NSA) is responsible for SIGINT, and the NSA collects SIGINT using signal communications between people, machines, and combinations of both. National Geospatial-Intelligence Agency (NGA) is responsible for the collection of IMINT using visual collection methods including cameras, radars, lasers, etc. MASINT is technically and scientifically based intelligence on unique attributes of the desired target. MASINT is the part of the Defense Intelligence Agency (DIA) and can be used to recognize, locate, and describe the target through nuclear, seismic, acoustic, material, radio frequency, and optical sciences. The Department of State (DOS), CIA, Department of Defense (DOD), Federal Bureau of Investigation (FBI), and other agencies use HUMINT sources, more commonly known as spies, to gather data via photography, document transfer, information download, etc. in enemy territory. OSINT relies on publicly available information, and a large synthesis of OSINT is performed at Foreign Broadcast Information Service (FBIS) and the National Air and Space Intelligence Center (NASIC). GEOINT belongs to the NGA and is generally collected from satellites and reconnaissance aircraft.

@@

\chapter[Appendix H: ????????]{?????????} \label{app:relations_table}

\section{???????????  Missing ??????????}

???????????  Missing ??????????

>>>>

DoS - Department of State

@@

Following the interception of SNM or weapon components, careful analysis of the composition, material structure, and relative sophistication of the physics package will allow forensic investigators to estimate the age of the device, the techniques which were utilized in order to create the physics package, and the most probable country of origin \cite{Glaser2008}. In the U.S., a number of organizations would be involved in these assessments. The National Technical Nuclear Forensics Center (NTNFC) is responsible for ensuring the nation's nuclear forensics capability.  The NTNFC is supported by the Department of Homeland Security (DHS), Department of Justice (DoJ), Federal Bureau of Investigation (FBI), Department of Defense (DoD), Department of Energy (DoE), Department of State (DoS), and the Office of the Director of National Intelligence (ODNI) \cite{DepartmentOfHomelandSecurit}.

@@

\ldots The State Department manages several programs that offer NIS scientists employment and research opportunities, so that they will be less likely to sell their knowledge to rogue nations. The International Science and Technology Center (ISTC) in Moscow and the Science and Technology Center in Ukraine (STCU), multilateral programs supported by the EU, Japan, and others, fund research in such areas as nuclear reactor safety, medicine, civil aviation, and energy production. A \enquote{Partners Program} links NIS scientists with contributing U.S. private sector companies, such as General Electric and Lockheed Martin, and academic institutes to assist their own R\&D efforts. The Civilian Research and Development Foundation (CRDF) is a private, non-profit foundation, originally funded by matching grants from DoD and George Soros. It supports U.S. collaboration and exchanges with former NIS defense scientists on civilian basic and applied research, and supports projects with commercial applications. The State Department also implements a pilot project for biological weapons scientists to \enquote{redirect} their expertise toward commercial endeavors. Among other activities, it has established a collaborative program between the Department of Agriculture's Agricultural Research Service and Russian scientists on agriculture research, and between the Department of Health and Human Services' Centers for Disease Control, National Institutes of Health, and Food and Drug Administration and Russian biomedical scientists to conduct research on infectious diseases. }

@@

Contributing Agency: DNDO, DTRA, DoS, Bechtel Nevada  \normalfont

@@

The lead U.S. agency responsible for implementing this policy will be the U.S. Department of State (DoS) supported by the U.S. Defense Threat Reduction Agency (DTRA). Preventing ISIL from acquiring a nuclear weapon, or at the very least SNM, falls under the purview of both the Bureau of Counterterrorism (CT) and the Bureau of International Security and Nonproliferation (ISN) within the DoS. Under this policy strategy, ISN is the principal actor in the DoS as they are the primary U.S. interface with the IAEA on the NPT \cite{Circular1970}. DTRA is a Combat Support Agency that supports the Combatant Commanders and Services in responding to any WMD threat. One aspect of this support is managing a research and development portfolio to develop tools and capabilities to detect and respond to WMD threats. DTRA is the agency within the U.S. government that works to verify treaties related to WMD are being carried out by all signatories, and thus will be able to support the DoS ISN in this policy \cite{DefenseThreatReductionAgency}.

@@

Lead Agency: DoT, DoD

@@

Lead Agency: DoS

@@

Lead Agency: DoS, DNDO/NTNFC

@@

The proposed policy  would strengthen the commitment to nuclear forensics by the NFAA through increased funding for the program from current \$20M 2015 budget to \$30M for 2016 \cite{Year2015}.  This money would be used for nuclear forensics-related research, to \enquote{provide steady funding streams} for nuclear engineers and physicists, to develop forensics capabilities in the area of post-detonation debris collection, and increase the number, fidelity, and participation in nuclear forensic exercises.  In addition to the increased funding, the proposed policy  would have the President, in coordination with the DoS and DoD, issue a declaratory policy regarding the U.S. response to the transfer of nuclear technology, materials, or weapons that result in a detonation of a nuclear weapon on the U.S. or one of its allies.  The proposed declaratory policy would be incorporated into the next NPR, but issued as a standalone document in the interim due to the dynamic ISIL threat.  The proposed policy wording is:

@@

\chapter{???????  Do final consistency checks for ISIS/ISIL, US/U.S., apostrophes, SLD/SLoD, quote marks  ???????????}

@@

Intelligence agencies gather information from physical and technical observations, conversations, searches, and human operations. These sources are grouped into Signals Intelligence (SIGINT), Imagery Intelligence (IMINT), Measurement and Signature Intelligence (MASINT), Human-Source Intelligence (HUMINT), Open-Source Intelligence (OSINT) and Geospatial Intelligence (GEOINT) \cite{Intelligen}. The National Security Agency (NSA) is responsible for SIGINT, and the NSA collects SIGINT using signal communications between people, machines, and combinations of both. National Geospatial-Intelligence Agency (NGA) is responsible for the collection of IMINT using visual collection methods including cameras, radars, lasers, etc. MASINT is technically and scientifically based intelligence on unique attributes of the desired target. MASINT is the part of the Defense Intelligence Agency (DIA) and can be used to recognize, locate, and describe the target through nuclear, seismic, acoustic, material, radio frequency, and optical sciences. The Department of State (DoS), CIA, Department of Defense (DoD), Federal Bureau of Investigation (FBI), and other agencies use HUMINT sources, more commonly known as spies, to gather data via photography, document transfer, information download, etc. in enemy territory. OSINT relies on publicly available information, and a large synthesis of OSINT is performed at Foreign Broadcast Information Service (FBIS) and the National Air and Space Intelligence Center (NASIC). GEOINT belongs to the NGA and is generally collected from satellites and reconnaissance aircraft.

@@

\chapter[Appendix H: Bilateral Relations between the U.S. \& Strategic Countries]{Bilateral Relations between the U.S. \&  Strategic Countries} \label{app:relations_table}

As a concise summary of U.S. foreign relations (as they relate to ISIL acquisition of nuclear weapons), a matrix of Bilateral Relations between the U.S. and Strategic Countries is provided in \autoref{tab:bilateral_table}.
\begin{table}[h]
\resizebox{1.38\textwidth}{!}{%
\begin{tabular}{|l|l|l|l|}
\multicolumn{1}{|c|}{Country}                               & \multicolumn{1}{c|}{Relationship with U.S.}                                                                                                                                                                                                                                                                                                                                                                                               & \multicolumn{1}{c|}{Military and Nuclear Capabilities}                                                                                                                                                                                                                                                                                                        & \multicolumn{1}{c|}{ISIL Position}                                                                                                                                                                                                                                                                                                                                                                      \\ \hline
Russia                                                      & \begin{tabular}[c]{@{}l@{}}Strained relationship due to economic sanctions, imposed for Russia's annexation of Crimea.\\ Predicted power struggle for new world order, and a historical rival.\\ Agreed to maintain dialogue in areas of terrorism, energy security, and\\ 			nuclear non-proliferation issues (Iran, North Korea).\end{tabular}                                                                                        & \begin{tabular}[c]{@{}l@{}}Military capability rank: 2 \cite{U.S.CentralIntelligenceAgency2012} (after the U.S.)\\ Defense spending: U.S.\$84.5 billion (2014) \cite{StockholmInternationalPeaceResearchInstitute2014}\\ Total NW inventory: 7,500; deployed strategic: 1,780 \cite{FederationofAmericanScientists2015}\end{tabular}                                                                                                                                               & \begin{tabular}[c]{@{}l@{}}More than 500 Russians are believed to have joined ISIL.\\ Russia considering supplying military equipment to Libya and Iraq. Agreed\\ 			to work together with  U.S. on facing ISIL.\end{tabular}                                                                                                                                                                             \\ \hline
China                                                       & \begin{tabular}[c]{@{}l@{}}China and the U.S. have a close economic partnership. \\ Two-way trade between China and the United States is \\ over U.S.\$590 billion in 2014. China is the second largest export country (after Canada).\end{tabular}                                                                                                                                                                                         & \begin{tabular}[c]{@{}l@{}}Military capability rank: 3 \cite{U.S.CentralIntelligenceAgency2012}\\ Defense Spending: U.S.\$145 billion (2014)\\ China has been increasing its annual military budget \\ over the last two decades. Total NW inventory: 250; deployed strategic: 0 \cite{StockholmInternationalPeaceResearchInstitute2014}\end{tabular}                                                                                            & \begin{tabular}[c]{@{}l@{}}Chinese Muslim Uyghur militants trying to flee the country and join ISIL\\ 			training camps in preparation for attacks back home.\end{tabular}                                                                                                                                                                                                                              \\ \hline
\begin{tabular}[c]{@{}l@{}}United\\ 			Kingdom\end{tabular} & \begin{tabular}[c]{@{}l@{}}Special relationship; close allies encompassing various issues, \\ including political and military, dating back to Churchillian era.\end{tabular}                                                                                                                                                                                                                                                           & \begin{tabular}[c]{@{}l@{}}Military capability rank: 5 \cite{U.S.CentralIntelligenceAgency2012} \\ Defense Spending: U.S.\$51.5 billion (2014) Talks of spending reduction.\\ Total NW inventory: 225; deployed strategic: 160 \cite{FederationofAmericanScientists2015}\end{tabular}                                                                                                                                              & \begin{tabular}[c]{@{}l@{}}The U.K. government declared ISIL a terrorist organization.\\ Hundreds of U.K. citizens are estimated to have joined ISIL.\\ 			U.K. citizens have been killed or kidnapped by ISIL.\end{tabular}                                                                                                                                                                            \\ \hline
France                                                      & \begin{tabular}[c]{@{}l@{}}France is the U.S.' oldest ally, relationship remains active and friendly. \\ 			France supported U.S. on various issues, but France was skeptical of U.S. invasion of Iraq.\end{tabular}                                                                                                                                                                                                                    & \begin{tabular}[c]{@{}l@{}}Military capability rank: 6 \cite{U.S.CentralIntelligenceAgency2012} \\ 			Defense spending: U.S.\$40 billion (2014)\\ 			Total NW inventory: 300; deployed strategic: 290 \cite{StockholmInternationalPeaceResearchInstitute2014}\end{tabular}                                                                                                                                                                       & \begin{tabular}[c]{@{}l@{}}France supported U.S.-led coalition against ISIL.\\ 			France's colonial history in Africa made it pivotal in the fight against\\ 			ISIL and Islamic terrorism in Africa. ISIL called for French\\ 			Muslim to join the group and perform terror attacks within France.\end{tabular}                                                                                         \\ \hline
\begin{tabular}[c]{@{}l@{}}North\\ 			Korea\end{tabular}    & \begin{tabular}[c]{@{}l@{}}North Korea has repeatedly threatened the U.S. with nuclear attacks.	\\  Relationship between the two countries are negative.		North Korea is not a party to NPT.\end{tabular}                                                                                                                                                                                                                               & \begin{tabular}[c]{@{}l@{}}Military capability rank: 36 \cite{U.S.CentralIntelligenceAgency2012} \\ 			Defense Spending: U.S.\$5.5 billion (2014) \cite{U.S.CentralIntelligenceAgency2012} \\ Total NW inventory: 6-8; deployed strategic: 0 \cite{StockholmInternationalPeaceResearchInstitute2014} \\ 			Recent reports suggest that North Korea may have up \\ to 20 nuclear warheads and could built as many as 100 by 2020 \cite{TheEditorialBoard2015a}.\end{tabular}                              & \begin{tabular}[c]{@{}l@{}}North Korea has publicly condemned U.S. effort in annihilating ISIL.\\ 			However, no concrete information has been gathered whether North\\ 			Korea and ISIL have made significant contact. U.S. worries North\\ 			Korea will sell nuclear technology to non-state actors.\end{tabular}                                                                                     \\ \hline
Israel                                                      & U.S. was the first country that that recognized independence of Israel in 1948.                                                                                                                                                                                                                                                                                                                                                         & \begin{tabular}[c]{@{}l@{}}Military capability rank: 11 \cite{U.S.CentralIntelligenceAgency2012}\\ 			Defense spending: U.S.\$17 billion (2014)\\ 			Total NW inventory: 80; deployed strategic: 0 \cite{StockholmInternationalPeaceResearchInstitute2014}.\end{tabular}                                                                                                                                                                          & \begin{tabular}[c]{@{}l@{}}Report from the UN reveals that the Israeli Defense Forces\\ (IDF) maintained regular contact with members of ISIL since May 2013.\end{tabular}                                                                                                                                                                                                                              \\ \hline
India                                                       & \begin{tabular}[c]{@{}l@{}} \enquote{Global strategic partnership} based on shared democratic values and\\ 			increasing convergence of interests on bilateral, regional and global issues. \\ Historically, India was close ally with Russia then\\ 			pioneered the Non-Aligned Movement (NAM).\end{tabular}                                                                                                                                   & \begin{tabular}[c]{@{}l@{}}Military capability rank: 4 \cite{U.S.CentralIntelligenceAgency2012}\\ 			Defense spending: U.S.\$38 billion (2014)\\ 			Total inventory: 90 - 100; deployed strategic: 0 \cite{StockholmInternationalPeaceResearchInstitute2014}.\end{tabular}                                                                                                                                                                       & \begin{tabular}[c]{@{}l@{}}ISIL identified Central Asia, including area north of India, as \enquote{a ripe recruitment area.} \\ 			The government considered ISIL as a periphery issue, but\\ 			ISIL's plan of expansion might include areas in India.\end{tabular}                                                                                                                                           \\ \hline
Pakistan                                                    & \begin{tabular}[c]{@{}l@{}}Over the decades, Pakistan has been a great ally to the U.S., but there\\ 			are instances where the two countries disagreed. U.S. views Pakistan \\ as its ally, but worries about the threat of terrorism in\\ 			Pakistan. Pakistan has received U.S.\$28.4 billion in military and non-military aid post 9/11.\end{tabular}                                                                                & \begin{tabular}[c]{@{}l@{}}Military capability rank: 17 \cite{U.S.CentralIntelligenceAgency2012}\\ 			Defense spending: U.S.\$7 billion (2014)\\ 			Pakistan is not a party to NPT.\\ 			Total NW inventory: Unknown.\end{tabular}                                                                                                                                                                & \begin{tabular}[c]{@{}l@{}}Many members of Pakistan's terrorist groups have joined ISIL. ISIL's leaked \\ secret memo spelled out the master plan on how to wage war\\ 			against the Pakistani military while trying to join forces with\\ 			local militants.\end{tabular}                                                                                                                            \\ \hline
Iran                                                        & \begin{tabular}[c]{@{}l@{}}Long and unpleasant history; current concerns about Iran are its\\ 			support for terrorism, pursuit of nuclear weapons, its opposition \\ towards Israelis and Arabs, and the country's poor human rights record.\end{tabular}                                                                                                                                                                             & \begin{tabular}[c]{@{}l@{}}Military capability rank: 23 \cite{U.S.CentralIntelligenceAgency2012}\\ 			Defense spending: U.S.\$6.3 billion (2014)\\ 			Currently, Iran and P5+1 countries are negotiating a nuclear deal to halt\\ 			Iran's nuclear development.\end{tabular}                                                                                                                    & \begin{tabular}[c]{@{}l@{}}Iran and ISIL are fighting for the crown of militant Islam. In this\\ 			case, our enemy's enemy is not our friend. The goal for the U.S. is to\\ 			defeat ISIL and to halt Iran's nuclear program.\end{tabular}                                                                                                                                                                  \\ \hline
Iraq                                                        & \begin{tabular}[c]{@{}l@{}}The Strategic Framework Agreement for a Relationship of Friendship and\\ 			Cooperation between the U.S. and the Republic of Iraq guides the\\ 			overall political, economic, cultural, and security ties with\\ 			Iraq.  This agreement is designed to help the Iraqi people stand\\ 			on their own and reinforce Iraqi sovereignty, while protecting 			U.S. interests in the Middle East.\end{tabular} & \begin{tabular}[c]{@{}l@{}}Military capability rank: 112 \cite{U.S.CentralIntelligenceAgency2012}\\ 			Defense spending: U.S.\$6 billion (2014)\\ 			The U.S. has been heavily investing in reforming and assisting the\\ 			country's military capabilities.\end{tabular}                                                                                                                          & \begin{tabular}[c]{@{}l@{}}ISIL is carving its territory from Iraq. The weak governance of Iraq\\ 			instigated ISIL's ambition. has been battling ISIL aided by a\\ 			coalition made up of Shiite militiamen and volunteers.\end{tabular}                                                                                                                                                             \\ \hline
Syria                                                       & \begin{tabular}[c]{@{}l@{}}The U.S.-Syria relationship is negative after having suspended diplomatic relations in 2014. \\ 			The U.S. has imposed economic sanction to Syria since 2004 due to\\ 			Syria's involvement in harboring terrorist groups. The U.S. is\\ 			deeply opposed to President Bashar al-Assad's regime,\\ 			specifically regarding chemical weapons use.\end{tabular}                                     & \begin{tabular}[c]{@{}l@{}}Military capability rank: 42 \cite{U.S.CentralIntelligenceAgency2012}\\ 			Defense spending: U.S.\$1.9 billion (2014)\\ Syrian's military hardware are mostly acquired from Russia and Iran.\\ 			Despite losing its military personnel, the Syrian military has been said\\ 			to become more flexible and capable especially in anti-guerrilla warfare.\end{tabular} & \begin{tabular}[c]{@{}l@{}}ISIL is gaining foothold in Syria, and carving its territory from the\\ 			country. Hence, Syrian government is fighting ISIL. The U.S. has\\ 			floated the idea to train the Syrian military opposition, and even\\ 			the Syrian military itself, in order to fight ISIL. Syria views\\ 			U.S.-led coalition against ISIL as a threat to their sovereignty.\end{tabular} \\ \hline
\caption{Matrix of Bilateral Relations between the U.S. and Strategic Countries}
\label{tab:bilateral_table}
